---
title: "Accuracy calculations"
author: "Robrecht Dockx"
format: pdf
editor: visual
---


```{r libraries, echo = FALSE, message = FALSE, comment= "", error = FALSE}
library(tidyverse)
library(terra)
library(git2rdata)
library(plotly)
library(sf)
library(INBOtheme)
library(assertthat)
library(caret)
library(rprojroot)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, 
                  comment = "", error = FALSE)

git_root <- rprojroot::find_root(rprojroot::is_git_root)
source(file.path(git_root, "source/scripts/nca_functions.R"))

data_folder <-  find_root_file("data",
                             criterion = has_file(
                               "histmap-validation.Rproj"))

```

# The data 

```{r}

data <- read.csv(paste0(data_folder, "/legende_attribuuttabel.xlsx - validatiepunten_timeseries.csv"), 
                        header = TRUE, encoding = "UTF-8")
cols <- c("X1850_true", 
          "X1904_tru", 
          "X1939_tru", 
          "X1981_tru", 
          "X1989_tru", 
          "X1774_true", 
          "X1873_true", 
          "X1969_true", 
          "ref")
new_cols <- c("1850", 
          "year_1904", 
          "year_1939", 
          "year_1981", 
          "year_1989", 
          "year_1774", 
          "year_1873", 
          "year_1969", 
          "year_2022")
data <- data %>%
  mutate(ref = as.integer(
    str_sub(
      as.character(
        LU_2022_1969_1873_1774_pred), start = 1, end = 1))) %>% 
  mutate(
  across(all_of(cols), 
         ~case_when(.x == 9 ~ "water", 
                    .x == 8 ~ "boomgaard",
                    .x == 7 ~ "moeras",
                    .x == 6 ~ "strand, slikke en schorre",
                    .x == 5 ~ "heide en duinen",
                    .x == 4 ~ "grasland",
                    .x == 3 ~ "bos",
                    .x == 2 ~ "bebouwing, tuin, infrastructuur",
                    .x == 1 ~ "akker")
           )) %>% 
  mutate(
    across(all_of(cols), as.factor)) %>% 
  rename_at(vars(cols),  ~ new_cols)

```

```{r}
# 2022 was set as reference and all other were compared with 2022
#

res1 <- calculate_accuracy(data$year_1774, data$year_2022)
res2 <- calculate_accuracy(data$year_1873, data$year_2022)
res3 <- calculate_accuracy(data$year_1969, data$year_2022)
```

```{r}
# Get UA and PA

data.frame(
  users = c(res1$byClass[, 5], res2$byClass[, 5], res3$byClass[, 5]),
  producers = c(res1$byClass[, 6], res2$byClass[, 6], res3$byClass[, 6]),
  lu = rep(str_remove(rownames(res1$byClass), "Class: "), 3),
  year = c(
    rep(1774, nrow(res1$byClass)),
    rep(1873, nrow(res2$byClass)), 
    rep(1969, nrow(res3$byClass))
  )
) %>%
  mutate(
    lu = as.factor(lu),
    year = as.factor(year)
  ) %>%
  ggplot() +
  geom_point(aes(x = users, y = producers, color = lu, shape = year)) +
  scale_color_discrete(name = "land use") +
  xlab("user's accuracy (precision)") +
  ylab("producer's accuracy (recall)") +
  theme_bw()

# Check for changes of the 3 other years 
observed_changes <- data.frame(
  map = as.factor(str_c(data$year_1774, data$year_1873,
                        data$year_1969,
                        sep = "-")),
  
  # ref = as.factor(str_c(data$year_2022, sep = "-")),
  mapchange1 = as.factor(ifelse(data$year_1774 == data$year_1873,
    "No change", "Change"
  )),
    mapchange2 = as.factor(ifelse(data$year_1873 == data$year_1969,
    "No change", "Change"
  )), 
    mapchange3 = as.factor(ifelse(data$year_1969 == data$year_2022,
    "No change", "Change"
  ))) %>%
  mutate(
    map = factor(map, levels = sort(unique(c(levels(map))))),
    map = factor(map, levels = levels(map))
  )



# Getting all possible combinations over the 3 years
observed_changes %>%
  pivot_longer(cols = 1:5, names_to = "type", values_to = "change") %>%
  group_by(change) %>%
  summarize(
    map = sum(type == "map"), 
  ) %>%
  ungroup()

```

# ReGeNesees

BEGIN exploration of other R packages to calculate the statistics       
instead of relying on hand-coded functions implementing Olofsson et al  
formulae                                                                
conclusion = the formulae are from a more general class of regression   
estimators; cf Sarndal model-assisted survey sampling                   
these can be implemented via the general R package for design-based     
inference "ReGeNesees"                                                  
also explored "mapac" R package, which also only implemented Olofsson   


```{r}

# install mapac package from gitlab
if (!c("mapac") %in% installed.packages()) {
  remotes::install_gitlab(
    repo = 'pflugmad/mapac',
    host = 'scm.cms.hu-berlin.de', quiet = F, force = T)
}
# install ReGenesees
if (!c("ReGenesees") %in% installed.packages()) {
  remotes::install_github("DiegoZardetto/ReGenesees")
}

```
function confusion_matrix() calculates the map-relevant version of error matrix
map-relevant means that each p_ij is weighted by area proportions for each stratum (map class) so this is different from default caret::confusionMatrix which was used in
function calculate_accuracy()
see https://pages.cms.hu-berlin.de/EOL/gcg_eo/06_accuracy_assessment.html

```{r}

```




